DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-02 paper/evaluation.tex

@@

However, the DeepPicar was shown to be capable of handling other potential WCETs. If given a larger 
deadline, the capabilities of DeepPicar as an autonomous vehicle platform would be greatly increased. 
For example, if tested with a deadline of 66.67 ms, or 15 fps, the platform would have passed several 
of the experiments performed. In actuality, the DeepPicar would have only been unsatisfactory in the 
4 model test where it would have missed the deadlines by $\sim$10 ms.

>>>>

\subsection{Benchmark Performance}
In order to determine how cache misses affects the performance of the DeepPicar, we test its ability by 
running synthetic benchmarks concurrently with the model. In each experiment, we run a single model on a 
number of cores, and a synthetic benchmark on the remaining idle cores. The results of each test can be 
seen in Table VII.

\begin{table}[h]
\centering
  \begin{tabular} {| l | l | l | l | l | l |}
  \hline
  \textbf{num cores} & 1 & 2 & 3 \\ \hline
  \textbf{mean} & 562.33 & 391.98 & 98.05 \\ \hline
  \textbf{L1 refs} & 3.16E+10 & 3.17E+10 & 3.06E+10 \\ \hline
  \textbf{L1 misses} & 5.41E+08 & 5.35E+08 & 4.98E+08 \\ \hline
  \textbf{L1 miss \%} & 1.71 & 1.69 & 1.62 \\ \hline
  \textbf{L2 refs} & 3.49E+09 & 3.23E+09 & 3.21E+09 \\ \hline
  \textbf{L2 misses} & 8.05E+08 & 7.54E+08 & 4.98E+08 \\ \hline
  \textbf{L2 miss \%} & 23.05 & 23.37 & 15.51 \\ 
  \hline
  \end{tabular}
  \caption{Real-time performance of the DeepPicar when benchmarks are run concurrently with the model.}
\end{table}

The presence of the benchmark(s) had a noticable effect on the DeepPicar, with all tests showing 
increases in inferencing operations. The least change happened when the model was run on 3 cores, and a 
single benchmark was run on the final core. Even with a slight increase in the number of L2 cache 
misses, inferencing was able to be completed in under 100 ms. The other tests, however, showed dramatic 
changes when the model was run on 1 and 2 cores. In both instances, L2 misses rise to $\sim$23\% of all 
references. The additional benchmark in the 1 core test was even more detrimental as the average 
inferencing time was at least 150 ms longer than when the model used 2 cores.

With the introduction of synthetic benchmarks, the DeepPicar failed to comply with the 50 ms WCET across 
all tests run. Even in the best case, inferencing operations take twice as long as the deadline to 
complete. As a result, if benchmarks were required to run during autonomous operation, the DeepPicar 
wouldn't be capable of meeting its deadlines.

@@

However, the DeepPicar was shown to be capable of handling other potential WCETs. If given a larger 
deadline, the capabilities of DeepPicar as an autonomous vehicle platform would be greatly increased. 
For example, if tested with a deadline of 66.67 ms, or 15 fps, the platform would have passed several 
of the experiments performed. In actuality, the DeepPicar would have only been unsatisfactory in the 
4 model test where it would have missed the deadlines by $\sim$10 ms, and the benchmark tests, where 
it would still miss by $\sim$33 ms at best.

### 2017-10-02 paper/intro.tex

% big problem
Executing these AI workloads on a embedded computing platform that can
be used in a car poses several additional challenges. First, many AI
workloads in vehicles are computationaly demanding and have strict
real-time requirements. For example, latency in a vision-based object
detection task may directly linked to safety of the vehicle. This
requires a high computing capacity as well as means to guaranteeing
the timings. On the other hand, the computing hardware platform must
also satisfy cost, size, weight, and power constraints, which require
highly efficient computing platform. These two conflicting
requirements  complicate the platform selection process as observed in
~\cite{Otterness2017}, which evaluated vision workloads on
NVIDIA's Tegra TX1 platform.
%% For example, while today's self-driving car
%% prototype equip more \$100,000 
%% of computers and sensors~\cite{juliussen2014emerging}, a study
%% found that aveage consumers are willing to pay much less amount of
%% extra cost for a self-driving capability~\cite{Daziano2017}.
%% https://qz.com/924212/what-it-really-costs-to-turn-a-car-into-a-self-driving-vehicle/

% related work and remaing problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While a real car
testbed would be most ideal, it is not only highly expensive but also
poses serious safety concerns that hinder development and exploration.
Therefore, there is a strong need for safer and less costly testbeds
that still employ state-of-the-art AI technologies. There are already
several relatively inexpensive RC-car based testbeds, such as MIT
RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
However, these RC-car testbeds still cost more than \$3,000.

% our goals
Instead, we want to build a low cost testbed that still employs the
state-of-the art AI technologies. Specifically, we focus on a deep
convolutional neural network (CNN) based end-to-end control system,
which was developed for a real self-driving car, NVIDIA
DAVE-2~\cite{Bojarski2016}, and use the same moethodology on a
smaller, safer and less costly setup. In developing the testbed, our
goals are (1) to analyze real-time issues in DNN based end-to-end
control; (2) to evalute real-time performance of contemporarly embedded
multicore platforms for the AI workloads.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research and education. From hardware perspective, 
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs state-of-the-art AI
techonologies, including a vision-based end-to-end control system that
utilizes a deep convolutional nueral network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}. 
The network has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is identifical to that of NVIDIA's DAVE-2
self-driving car~\cite{Bojarski2016}, which uses much more powerful
computer (Drive PX computer~\cite{drivepx}) than Raspberry Pi~3.
We chose to use Pi 3 not only because of its
affordability but also because it is representative
of today's mainstream low-end embedded multicore platforms, found in
smartphones and other embedded devices.

@@

% how we trained (maybe moved to a later section)
We apply a standard imitation learning methodology to train the cat to
follow tracks on the ground. We collect data for 
taining and validation by manually  
controlling the RC car and recording the vision (from the webcam
mounted on the RC-car) and the human control inputs. We then train the
network offline using the collected data on a desktop computer, which
equips a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
back to the Raspberry Pi, which is then used to perform interference
operations---locally on the Pi---in the car's main control loop in
real-time. For real-time control, each inference operation must
be completed within the desired control period. (e.g., 50ms period for
20Hz control frequency.)
% how we evaluated (in terms real-time performance)

@@

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present the design and implementation of a
    low-cost autonomous vehicle testbed DeepPicar, which utilizes the
    state-of-the-art aritificial intelligence techniques.
  \item We provide an analysis and case-study of real-time issues in the
    DeepPicar.
  \item We systematically compare real-time computing capabilities of
    multiple embedded computing platforms in the contex of a
    vision-based autnomous driving.
\end{itemize}

>>>>

% big problem
Executing these AI workloads on a embedded computing platform that can
be used in a car poses several additional challenges. First, many AI
workloads in vehicles are computationally demanding and have strict
real-time requirements. For example, latency in a vision-based object
detection task may directly linked to safety of the vehicle. This
requires a high computing capacity as well as means to guaranteeing
the timings. On the other hand, the computing hardware platform must
also satisfy cost, size, weight, and power constraints, which require
highly efficient computing platform. These two conflicting
requirements  complicate the platform selection process as observed in
~\cite{Otterness2017}, which evaluated vision workloads on
NVIDIA's Tegra TX1 platform.

%% For example, while today's self-driving car
%% prototype equip more \$100,000
%% of computers and sensors~\cite{juliussen2014emerging}, a study
%% found that aveage consumers are willing to pay much less amount of
%% extra cost for a self-driving capability~\cite{Daziano2017}.
%% https://qz.com/924212/what-it-really-costs-to-turn-a-car-into-a-self-driving-vehicle/

% related work and remaining problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While a real car
testbed would be most ideal, it is not only highly expensive but also
poses serious safety concerns that hinder development and exploration.
Therefore, there is a strong need for safer and less costly testbeds
that still employ state-of-the-art AI technologies. There are already
several relatively inexpensive RC-car based testbeds, such as MIT
RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
However, these RC-car testbeds still cost more than \$3,000.

% our goals
Instead, we want to build a low cost testbed that still employs the
state-of-the art AI technologies. Specifically, we focus on a deep
convolutional neural network (CNN) based end-to-end control system,
which was developed for a real self-driving car, NVIDIA
DAVE-2~\cite{Bojarski2016}, and use the same methodology on a
smaller, safer and less costly setup. In developing the testbed, our
goals are (1) to analyze real-time issues in DNN based end-to-end
control; (2) to evaluate real-time performance of contemporary embedded
multicore platforms for the AI workloads.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research and education. From hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs state-of-the-art AI
technologies, including a vision-based end-to-end control system that
utilizes a deep convolutional neural network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The network has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is identical to that of NVIDIA's DAVE-2
self-driving car~\cite{Bojarski2016}, which uses much more powerful
computer (Drive PX computer~\cite{drivepx}) than Raspberry Pi~3.
We chose to use Pi 3 not only because of its
affordability but also because it is representative
of today's mainstream low-end embedded multicore platforms, found in
smartphones and other embedded devices.

@@

% how we trained (maybe moved to a later section)
We apply a standard imitation learning methodology to train the cat to
follow tracks on the ground. We collect data for
training and validation by manually
controlling the RC car and recording the vision (from the webcam
mounted on the RC-car) and the human control inputs. We then train the
network offline using the collected data on a desktop computer, which
equips a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
back to the Raspberry Pi, which is then used to perform interference
operations---locally on the Pi---in the car's main control loop in
real-time. For real-time control, each inference operation must
be completed within the desired control period. (e.g., 50ms period for
20Hz control frequency.)
% how we evaluated (in terms real-time performance)

@@

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present the design and implementation of a
    low-cost autonomous vehicle testbed DeepPicar, which utilizes the
    state-of-the-art artificial intelligence techniques.
  \item We provide an analysis and case-study of real-time issues in the
    DeepPicar.
  \item We systematically compare real-time computing capabilities of
    multiple embedded computing platforms in the context of a
    vision-based autonomous driving.
\end{itemize}
